---
title: "Buffer_Protocol_Rough_Draft"
author: "Sofia Ingersoll"
format: html
editor: visual
---

*Rough Draft pls no judging yet*

# Protocol for Creating Buffer Zones & Selecting Random Addresses in Buffer Regions

The following code describes a step-by-step process for creating a map containing buffer zones around points of interest. In this documentation, the data utilized is the [US Wind Data](https://dataverse.harvard.edu/file.xhtml?fileId=7339850&version=1.0), this data is associated with the "Replication Data for: Prevalence and predictors of wind energy opposition in North America", <https://doi.org/10.7910/DVN/LE2V0R>, Harvard Dataverse, V1, 2023. The collaborators on that project include: Stokes, Leah; Franzblau, Emma; Lovering, Jessica R.; Miljanich, Chris.

\~ include more about what the data is about and the outcomes of making visualization \~

### Loading Libraries

```{r}
library(tidyverse)        # essential r package 
library(tmap)             # package aids in the creation of buffer zones
library(sf)               # package simplifies spatial dataframes
                          # by providing 'sticky geometry'
library(smoothr)          # aesthetic and visual aid for buffers created
library(sf)
library(spData)
library(ggplot2)
library(raster)
```

### Read in the Data

```{r}
wind_data <- read.csv("data/wind_data_usa.csv")     # reading in & storing data
```

#### Confirm the Data Loaded Properly

```{r}
head(wind_data)                  # displays the first 6 rows of the data
                                 # along with all of the columns 
```

## Wrangling & Subsetting

*want:*

*- a scalelable map of USA*

*- the name of the plants as points*

*- adjustable buffer zones (donuts) around the plants*

*- ability to randomly select addresses within the buffer*

*- filter for specific groups of interest*

### **Need to convert lat/long into raster data or geometry column**

**Below is pseudocode**

```{r}
rasterFromXYZ(wind_data, crs = "4326")
```

```{r}
coordinates(wind_data.sp) <- c('longitude', 'latitude')

```

### **Need to set CRS: likely EPSG:4326**

```{r}
st_transform(wind_data, crs = '4326')
```

### Subset of plant names 

I have a strong inclination that I won't need this using tmaps, but just in case for now.

```{r}

```

## Initial Visualization of the Data

**More psuedocode**

```{r}
tmap_mode('view')

tm_shape(wind_data, coords =  c("longitude", "latitude")) +
  tm_polygons() +
  tm_symbols('plant_name')
```

```{r}

mapview(wind_data, xcol = "longitude", ycol = "latitude", crs = 4269, grid = FALSE)
```

## Buffers

Buffers create polygons representing a set distance from a feature.

**Below is psuedocode for now until the cleaning is configured right**

```{r include=TRUE}
wind_data_buffer <- st_buffer(wind_data, dist = 5000)

tm_shape(wind_data_buffer) +
  tm_polygons()
```
