---
title: "Buffer_Protocol_Rough_Draft"
author: "Sofia Ingersoll"
format: html
editor: visual
---

*Rough Draft pls no judging yet*

# Protocol for Creating Buffer Zones & Selecting Random Addresses in Buffer Regions

The following code describes a step-by-step process for creating a map containing buffer zones around points of interest. In this documentation, the data utilized is the [US Wind Data](https://dataverse.harvard.edu/file.xhtml?fileId=7339850&version=1.0), this data is associated with the "Replication Data for: Prevalence and predictors of wind energy opposition in North America", <https://doi.org/10.7910/DVN/LE2V0R>, Harvard Dataverse, V1, 2023. The collaborators on that project include: Stokes, Leah; Franzblau, Emma; Lovering, Jessica R.; Miljanich, Chris.

\~ include more about what the data is about and the outcomes of making visualization \~

Analysis of these areas will provide insight into local resistance and spatially distorted signalling in relation to wind power infrastructure and climate policy.

### Loading Libraries

The following libraries were selected based on their functionality and ability to optimize our data for mapping.

```{r}
# Loading Libraries
library(tidyverse)        # essential r package 
library(sf)               # package simplifies spatial dataframes
library(raster)
library(tmap)
library(terra)
library(stars)
library(smoothr)          # aesthetic and visual aid for buffers created
```

### Read in the Data

To simplify the following step, it is important to organize your folders in a way that makes sense for your workflow. In many cases, the data sets we work with are typically too large to be uploaded to GitHub. As a result, a common practice is storing your data in a folder, directly outside of your repository in a folder labeled "data".

The code chunk below for `read.csv` demonstrates how to exit your current location using `..` and enter the desired folder location using `/`. It is important that your file path does not contain any spaces and is directly reflective of the file path for the data you wish to read in.

```{r}
wind_data <- read.csv("../data/wind_data/wind_data_usa.csv")  
# reading in & storing data
```

##### EJ Screen Data Potentially Use for Base Map

```{r include=TRUE, eval=FALSE, warning=FALSE, error=FALSE}
ejscreen <- st_read("../data/EJSCREEN_2023_BG_StatePct_with_AS_CNMI_GU_VI.gdb",
                    quiet = TRUE)
```

##### Confirm the Data Loaded Properly

```{r}
head(wind_data)                  # displays the first 6 rows of the data
                                 # along with all of the columns 
```

## Wrangling & Subsetting

*want:*

*- a scaleable map of USA w/ state boundaries*

*- the name of the plants as points*

*- adjustable buffer zones (donuts) around the plants*

*- ability to randomly select addresses within the buffer*

*- filter for specific groups of interest*

### **Converting lat/long into Raster Data (i.e. sticky geometries)**

Below we will use the package `sf` to convert the lat/long data into a raster geometry column. In this single line, we will also be assigning the CRS EPSG:4326 to the sf data frame. The CRS was selected because it provides a relatively proportionate display of the United States. We are open to suggestions regarding our CRS if a different project better fits our data.

```{r}
wind_sf <- wind_data %>%             # calls desired dataset
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326) 
                                     # creates geometry column with desired crs 
```

### Check-in

Were the lat/long columns correctly converted into a geometry column? `setdiff()` is a way to quickly determine the differences between two data sets.

```{r}
# seeing what is different
setdiff(wind_data, wind_sf)
```

```{r}
# checking column names
print(colnames(wind_data))

print(colnames(wind_sf))
```

#### **Confirm CRS is correct**

```{r}
glimpse(crs(wind_sf))                  # output should reveal WGS84, EPSG:4326
```

## Initial Visualization of the Data

*ggplot psuedocode*

### Creating a Base Map Layer

Using the world dataset provided in the sf package, we will filter out the United States

*First attempt is not the best method because it does not include the states & their borders*

```{r}
#us <- world %>%                              # calls sf dataset
 # filter(name_long == "United States") %>%   # filters for states in the US
  #st_transform(us, crs = "EPSG:4326")        # converts the CRS to match wind_sf

#us_map <- ggplot(us) +
 # geom_sf(color = 'black')

#us_map
```

*Second attempt is missing Alaska and Hawaii but has borders*

```{r}
#usa <- st_as_sf(maps::map('state', fill = TRUE, plot = FALSE))

#ggplot(usa) +
 # geom_sf(color = 'black',
  #        fill = 'white',
   #       size = 0.125) +
  #coord_sf(crs = st_crs(4326))
```

*Third Attempt*

```{r}
#str(us_map())

#us_states_for_mapping <- us_map(
 # regions = "states")

#ggplot(us_states_for_mapping) +
 # geom_sf(color = "black")
```

*Preliminary Plot of Wind Data*

```{r}
ggplot() +
  geom_sf(data = wind_sf) 
```

*Fourth Attempt but EJ Screen Data*

The issue with the ejscreen data are the inclusion of US territories found in region 9. Below I am working on a process to filter out those specific locations from the data without removing desired states in the Pacific region.

```{r}
ggplot() +
  geom_sf(data = ejscreen)
  geom_sf(data = wind_sf) 
```

#### Regions of Interest

[Region Codes](https://ncsesdata.nsf.gov/docs/location.html) reveals the various regions throughout the US that are identifiable using codes between 1:10. US Territories are categorized under the Pacific Region Code, 9. Other states like CA, OR, HA, AK, and WA are also in the Pacific Category.

```{r}
unique(ejscreen$REGION)
```

#### Identifying the State Names for US Territories

To properly filter out these unwanted observations, it's best to identify the exact names of the unwanted regions. We can do this using the `unique` function on the state_name column.

```{r}
print(unique(ejscreen$STATE_NAME))
```

#### Removing Regions Outside of Interest

**not subsetting properly, maybe use** `subset()`

```{r}
us_basemap_data <- ejscreen %>% 
  filter(STATE_NAME != c('Puerto Rico', 'Guam', 'Virgin Islands', 'Northern Mariana IS', 'American Samoa')) %>% 
  filter(REGION != 10)
```

#### Next Idea: Create a Mask using US Country Coordinates

[Geojson Coordinate Mapper for Masks](http://geojson.io/#map=2.59/42.2/-96.48)

```{r}

```

*Preliminary Plot of Filtered EJScreen*

```{r}
ggplot() +
  geom_sf(data = us_basemap_data) +
  geom_sf(data = wind_sf) +
  theme_bw()
```

**improperly subset, so unable to map**

```{r}
tmap_mode('view')

tm_shape(us_basemap_data) +
  tm_shape(wind_sf) 

```

```{r}
ggplot() +
  geom_sf(data = us_basemap_data) +
  geom_sf(data = wind_sf) +
  theme_bw() +
  labs(title = "US Wind Infrastructure Plants") 
 # annotation_scale(plot_unit = "km") + # add scale bar
  #annotation_north_arrow( # add north arrorw
   # location = "tr",
    #pad_x = unit(0.2, "in"),
    #pad_y = unit(0.2, "in"),
    #style = ggspatial::north_arrow_nautical( # customize north arrow
     # fill = c("grey40", "white"),
    #  line_col = "grey20"
    #)
#  )

```

## Subsetting Wind Plant Locations in US

\*\* double check written description here \*\*

The code below selects only coordinates that intersect with the wind data and us mapping information

```{r}
# wind_sf <- wind_sf[us_basemap_data,]       # creates a combined subset of the                                                   wind data and the us}
```

## Buffers

Buffers create polygons representing a set distance from a feature.

The buffer zone dimensions were selected to correlate with the research presented in "Replication Data for: Prevalence and predictors of wind energy opposition in North America".

**Below is psuedocode for now until the cleaning is configured right**

```{r include=TRUE}
wind_data_buffer <- st_buffer(wind_sf, dist = 3000)
```

### Final Base Map of US

-   Need to include state lines for greater understanding and accuracy for collecting addresses

```{r}

```

**tmap psuedocode**

```{r}
tmap_mode('view')

tm_shape(wind_sf) +
  tm_polygons() +
  tm_symbols('plant_name')
```

```{r include=TRUE}
tm_shape(wind_data_buffer) +
  tm_polygons()
```
