---
title: "Buffer_Protocol_Rough_Draft"
author: "Sofia Ingersoll"
format: html
editor: visual
---

*Rough Draft pls no judging yet*

# Protocol for Creating Buffer Zones & Selecting Random Addresses in Buffer Regions

The following code describes a step-by-step process for creating a map containing buffer zones around points of interest. In this documentation, the data utilized is the [US Wind Data](https://dataverse.harvard.edu/file.xhtml?fileId=7339850&version=1.0), this data is associated with the "Replication Data for: Prevalence and predictors of wind energy opposition in North America", <https://doi.org/10.7910/DVN/LE2V0R>, Harvard Dataverse, V1, 2023. The collaborators on that project include: Stokes, Leah; Franzblau, Emma; Lovering, Jessica R.; Miljanich, Chris.

\~ include more about what the data is about and the outcomes of making visualization \~

### Loading Libraries

```{r}
library(tidyverse)        # essential r package 
library(tmap)             # package aids in the creation of buffer zones
library(sf)               # package simplifies spatial dataframes
                          # by providing 'sticky geometry'
library(smoothr)          # aesthetic and visual aid for buffers created
library(sf)               # sticky geometries
library(spData)
library(ggplot2)
library(raster)
```

### Read in the Data

```{r}
wind_data <- read.csv("data/wind_data_usa.csv")     # reading in & storing data
```

#### Confirm the Data Loaded Properly

```{r}
head(wind_data)                  # displays the first 6 rows of the data
                                 # along with all of the columns 
```

## Wrangling & Subsetting

*want:*

*- a scaleable map of USA w/ state boundaries*

*- the name of the plants as points*

*- adjustable buffer zones (donuts) around the plants*

*- ability to randomly select addresses within the buffer*

*- filter for specific groups of interest*

### **Converting lonvert lat/long into Raster Data (i.e. sticky geometries)**

Below we will use the package sf to convert the lat/long data into a raster geometry column. In this single line, we will also be assigning the CRS EPSG:4326 to the sf data frame. The CRS was selected because it provides a relatively proportionate display of the United States. We are open to suggestions regarding our CRS if a different project better fits our data.

```{r}
wind_sf <- wind_data %>%             # calls desired dataset
  st_as_sf(coords = c("longitude", "latitude"), crs = 4326) 
                                     # creates geometry column with desired crs 
```

### **Confirm CRS is correct**

```{r}
crs(wind_sf)                         # output should reveal WGS84, EPSG:4326
```

## Initial Visualization of the Data

**ggplot psuedocode**

### Creating a Base Map Layer

Using the world dataset provided in the sf package, we will filter out the United States

```{r}
us <- world %>%                              # calls sf dataset
  filter(name_long == "United States") %>%   # filters for states in the US
  st_transform(us, crs = "EPSG:4326")        # converts the CRS to match wind_sf
```

#### Subsetting Wind Plant Locations in US

\*\* double check written description here \*\*

The code below selects only coordinates that intersect with the wind data and us mapping information

```{r}
wind_sf <- wind_sf[us,]                      # creates a combined subset of the                                                   wind data and the us
```

#### Creating a Grid of Equal Area Polygons for Plant Plotting

\*\* idea not necessarily necessary \*\*

#### US Base map plotted

-   Need to include state lines for greater understanding and accuracy for collecting addresses

```{r}
us_map <- ggplot(us) +
  geom_sf(color = 'black')

us_map
```

```{r}

```

**tmap psuedocode**

```{r}
tmap_mode('view')

tm_shape(wind_sf) +
  tm_polygons() +
  tm_symbols('plant_name')
```

## Buffers

Buffers create polygons representing a set distance from a feature.

**Below is psuedocode for now until the cleaning is configured right**

```{r include=TRUE}
wind_data_buffer <- st_buffer(wind_data, dist = 5000)

tm_shape(wind_data_buffer) +
  tm_polygons()
```
